---

title: "Reproducible Research: Peer Assessment 1"
output: 
html_document:
keep_md: true

---

## Loading the data

### Code to load the data:

```{r}
data <- read.csv("activity.csv", header=T, sep=",")
```

### A sample of the data loaded:

```{r}
head(data)

```

## What is mean total number of steps taken per day?

```{r}
data.stepsDay <- aggregate(data$steps, by=list(data$date), FUN=sum, na.rm=F)
data.stepsDayMean <- mean(data.stepsDay$x, na.rm=T)
data.stepsDayMedian <- median(data.stepsDay$x, na.rm=T)


hist(data.stepsDay$x, 
     breaks=30,
     main="Steps taken per day", 
     col="green", 
     xlab="Steps")


```

## What is the average daily activity pattern?
### Wasn't the clearest of questions, assumed the following"
We want to map the steps taken per interval? 

```{r}

data.stepsInterval <- aggregate(data$steps, by=list(data$interval), FUN=mean, na.rm=T)

plot( x=data.stepsInterval[,1], 
      y=data.stepsInterval[,2], 
      type="l",
      col="green",
      main="Mean of steps per interval",
      ylab="Steps", 
      xlab="Interval")

```

## Inputing missing values
```{r}

nas   <- sum(is.na(data))
compl <- sum(complete.cases(data))

data.clean <- cbind(data, data.stepsInterval[,2])
names(data.clean)[4] <- c("mean")

data.clean$steps <- ifelse( is.na(data.clean$steps), data.clean$mean, data.clean$steps)
data.clean.stepsDay <- aggregate(data.clean$steps, by=list(data.clean$date), FUN=sum)
data.clean.stepsDayMean <- mean(data.clean.stepsDay$x)
data.clean.stepsDayMedian <- median(data.clean.stepsDay$x,)

hist(data.clean.stepsDay$x, 
     breaks=20,
     main="NEW Steps per DAy", 
     col="green", 
     xlab="Steps")




```

## Are there differences in activity patterns between weekdays and weekends? 

There are differences but the overall curve looks relatively similar, albeigt slightly lowered in volume.

```{r}

data.clean$date <- strptime(data.clean$date, "%Y-%m-%d")
data.clean$weekend <- (weekdays(data.clean$date) %in% c("Sunday", "Saturday"))
data.weekend <- data.clean[data.clean$weekend == TRUE,]
data.weekday <- data.clean[data.clean$weekend == FALSE,]

data.weekend.steps <- aggregate(data.weekend$steps, by=list(data.weekend$interval), FUN=mean)
data.weekday.steps <- aggregate(data.weekday$steps, by=list(data.weekday$interval), FUN=mean)

par(crow=c(2,0))

plot( x=data.weekend.steps[,1], 
      y=data.weekend.steps[,2], 
      type="l",
      col=124,
      main="Weekend",
      ylab="Steps", 
      xlab="Interval")

plot( x=data.weekday.steps[,1], 
      y=data.weekday.steps[,2], 
      type="l",
      col=554,
      main="Weekday",
      ylab="Steps", 
      xlab="Interval")


```


